---
title: "Assignment 3"
author: "James Naphtali"
date: "November 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Import data
```{r R.options=list(max.print=10)}
count_data <- read.csv("C:/Users/james/Documents/R Analysis/Bio720/Assignment 3/eXpress_dm_counts.csv", row.names = 1)
head(count_data)
```
##Summary of data
```{r R.options=list(max.print=12)}
str(count_data)
head(summary(count_data))
```

## Question 2
```{r count_data}
output <- vector("double", ncol(count_data))
col_means <- function(x, transformLog2) {
  
  for (i in seq_along(x)) {
    if (transformLog2 == TRUE) {
		output[[i]] <- mean(log2(x[[i]]+0.1)) ##Add 0.1 to default expression values to avoid non-values after log2. 
  } else
    output[[i]] <- mean(x[[i]])
  }
  output
  
}
  
log_means <- col_means(count_data, TRUE)
log_means
means <- col_means(count_data, FALSE)
means
```
## Question 3
```{r}
names(means) <- colnames(count_data)
means
means[which.min(means)]
means[which.max((means))]
names(log_means) <- colnames(count_data)
log_means
log_means[which.min(log_means)]
log_means[which.max((log_means))]
```
The lowest and highest expression values in non-log transformed dataset are small females' thoracic horns and large males' wings. In the log transformed dataset, large females' wings are the lowest   
in expression value and the highest are small females' wings. 
## Question 4
```{r}
means_sapply <- sapply(count_data, mean)
means_sapply
system.time(sapply(count_data, mean))#System time using the sapply()
system.time(col_means(count_data, transformLog2 = TRUE)) #System time using my custom function
```
Using the sapply() function, R calculates the mean quicker than using the user defined for loop function. 

##Question 5
```{r}
mean_rish <- colMeans(count_data)
mean_rish
system.time(colMeans(count_data))
```
##Question 6
```{r}
##Objective - loop over rows and obtain row means.
output <- vector("double", nrow(count_data)) #Change output vector to number of rows
row_means <- function(x, transformLog2) {
  
  for (i in seq_along(x)) {
    if (transformLog2 == TRUE) {
		output[[i]] <- mean(log2(x[[i]]+0.1)) ##Add 0.1 to default expression values to avoid non-values after log2. 
  } else
    output[[i]] <- mean(x[[i]])
  }
  output
  
}
mean_gene_exp<- row_means(count_data, TRUE)
names(mean_gene_exp) <- rownames(count_data)
head(mean_gene_exp)
```
##Question 7
```{r}
##Objective - subset dataframe to column names containing all colnames with the following strings: "lg_male_hdhorn" and "sm_male_hdhorn"
all_male_grep <- grep("_male_hdhorn", names(count_data), value = TRUE)
male_hdhorn_df <- count_data[, all_male_grep] #Subset dataframe with grepped column name vector
male_mean_gene <- rowMeans(male_hdhorn_df)
##Subset dataframe with only large males
lg_male_grep <- grep("lg_male_hdhorn", names(count_data), value = TRUE)
lg_male_hdhorn_df <- count_data[, lg_male_grep] #Subset dataframe with grepped column name vector
lg_mean_gene <- rowMeans(lg_male_hdhorn_df)
##Subset dataframe with only small males
sm_male_grep <- grep("sm_male_hdhorn", names(count_data), value = TRUE)
sm_male_hdhorn_df <- count_data[, sm_male_grep] #Subset dataframe with grepped column name vector
sm_mean_gene <- rowMeans(sm_male_hdhorn_df)
diff_means <- abs(lg_mean_gene - sm_mean_gene) #Calculate difference in absolute mean gene expression between large and small males head horns. 
head(diff_means)
```
##Question 8
```{r}
plot(male_mean_gene, diff_means, main="Mean Gene and Difference in Mean Gene Expression", xlab="Mean Gene Expression", ylab="Difference in Mean Gene Expression") #Plot non-log transformed means vs. difference of means of male head horns
log_gene_male_means <- log2(male_mean_gene)
log_gene_means_diff <- log2(diff_means)
plot(log_gene_male_means, log_gene_means_diff, main="Log2 Transformed Mean Gene and Difference in Mean Gene Expression", xlab="Log 2 Mean Gene Expression", ylab="Log 2 of Difference in Mean Gene Expression")
```