---
title: "ID_Bio720_Week6"
author: "James Naphtali"
date: "October 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Question 1
A HTML and a .RMD file were generated. 

##Question 2
R is a high level language since it has many layers of abstraction over machine code. High-level language is more human-readable compared to low-level. 

##Question 3
R is an interpretive language. Interpretive languages does not have to be compiled into another language to run, compared to C++, a compiled language, where the code is converted into machine-code to be run

##Question 4
R is slow due to its interactive (run by line) nature, also its high number of built-in functions to be loaded. R was designed for stability to prevent errors and consistency during statistical analysis 

##Question 5
C and Fortran. Fortran and C are faster and more stable languages than R for core calculations.

##Question 6
asd

##Question 7
Logical (Boolean), integer, real (double or floating point), complex, string, and raw

##Question 8
typeof(x) classifies datatypes based on how R classifies atomic types, while mode(x) shows what type it stors the data.  

##Question 9
y <- c(3,4,5) is considered a double datatype, while z=3:5 is considered an integer datatype in R by using typeof(x) function.

##Question 10
R predicts that there will probably be fractions from operations such as division and square roots and treats the data as double. The square operation results in a double for flexibility in handling large input values as well as floating-point data. 

##Question 11
1. integer
2. character
3. character
4. logical
5. character

##Question 12
We can coerce a variable using the == operator, telling R to equal x (an double) to y (a character) by converting x into a character. This is different from the = sign which is used to denote variables and literally means "is". We can also use functions such as as.character() to coerce variables into characters respectively. We can coerce a to b by using is.logical(). 

##Question 13
Use the function integer(), specifically integer(10). It outputs 10x zeroes as we did not define any other arguments, which results in the function outputting its default parameters. 

##Question 14
Sum is 20, as falses have a numeric value of 0.

##Question 15
We convert the objects in the environment into a list of characters, and using the remove function, remove the list. 
The objects are numeric double types.

```{r}
genotype <- c("wildtype","wildtype","wildtype","mutant","mutant","mutant")
str(genotype)
genotype2 <- (rep(c("wildtype", "mutant"), each=3))
str(genotype2)
genotype3 <- gl(n = 2, k = 3, labels = c("wildtype", "mutant"))
str(genotype3)
```
####Genotype and genotype 2 are the same data type, a vector of characters. Genotype 3 is generated as a factor with 2 levels. 

##Question 16
```{r}
genotype2_factor=factor(genotype2)
str(genotype2_factor) 
```
####we can check if it is a factor using str(). Specifying genotype2 as a factor afterwards yields the same datatype (a factor with 2 levels) as genotype3 although the order between factorized genotype2 is different than genotype3   

##Question 17
```{r}
day <- gl(n = 2, k = 1 , length = 6, labels = c(3, 6)) #using the same code as genotype3
str(day)
```

##Question 18
```{r}
as.character(day)
```

##Question 19
```{r}
as.numeric(day) #made numerics of factor levels
```

##Question 20
```{r}
as.numeric(as.character(day)) #convert to characters first and then coerce the correct lengths in character to numerics. 
```

##Question 21
```{r}
gene1 <- c(3, 4, 7, 9, 12, 6)
gene2 <- c(11, 17, 12, 25, 23, 7)
gene3 <- c(100, 103, 97, 94, 106, 111)

gene_mat1 <- cbind(gene1, gene2, gene3)
gene_mat1

gene_mat2 <- matrix(c(gene1, gene2, gene3), nrow = 6, ncol = 3, byrow=FALSE)
gene_mat2

identical(gene_mat1, gene_mat2)
```
####difference between using cbind() and matrix () is only that the column names are preserved using cbind. 

##Question 22




