---
title: "Bio720 Simulation Assignment"
author: "James Naphtali"
date: "December 5, 2018"
output: html_document
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```  

##Question 2  

```{r}
diploid_selection <- function(p0 = 0.01, w_AA = 0.8, w_aa = 0.16, w_Aa = 0.7, n = 100) {
    
    # Initialize vectors to store allele frequencies and mean pop fitness
    p <- rep(NA, n) 
    
    w_bar <- rep(NA, n) 
     
    # starting conditions
    p[1] <- p0 # starting allele frequency have to start the counter/for loop somewhere, so assign input of p0 with 0.01 to start of index 1 in p list 

    w_bar[1] <- (((p[1]^2)) * w_AA) + (2*p[1]*(1-p[1])*w_Aa) + ((((1-p[1])^2) * w_aa))
    
    # now we need to loop from generation to generation
    for (i in 2:n) {
        w_bar[i-1] <- (((p[i-1]^2)) * w_AA) + (2*p[i-1]*(1-p[i-1])*w_Aa) + ((((1-p[i-1])^2) * w_aa)) #Mean fitness 
        p[i] <- ((p[i-1]^2) * (w_AA/w_bar[i-1])) + ((p[i-1] * (1-p[i-1])) * (w_Aa/w_bar[i-1])) #Allele frequency of A
        
    }
    
    return(p)
}
#Graph model 
p <- diploid_selection(p0 = 0.01, w_AA = 0.7, w_aa = 0.3, w_Aa = 0.7, n =100)
generations <- 1:length(p)
plot(p ~ generations, pch =20,
     ylab = "allele frequency",
     xlab = "generation")
```  

##Question 3  
```{r}
gen_drift_sim <- function(indiv=1000, p = 0.4, n=1000){
  num_alleles = indiv * 2
  p_vec <- rep(NA, n)
  p_vec[1] <- p
  
  plot(1,0, type="n", xlim=c(1,n), ylim=c(0,1), xlab="Generation", ylab="Frequency of allele A")
  
  for (i in 2:n){
    allele_counts <- sample(c("A", "a"), size = num_alleles, replace = TRUE, prob = c(p, (1-p))) 
    p <- length(allele_counts[allele_counts == "A"])/(num_alleles)
    p_vec[i] <- p
    
    
    points(i, p, pch=19, col="black")
    
  }
}
gen_drift_sim()
```  

##Question 4 
```{r}
gen_drift_sim2 <- function(indiv=200, p0 = 0.5, n=100, num_sim = 1000){
  num_alleles = indiv * 2
  p_vec <- rep(1, n)
  
  
  for (x in 1:num_sim){
    p <- p0
  
    for (i in 2:n){
      allele_counts <- sample(c("A", "a"), size = num_alleles, replace = TRUE, prob = c(p, (1-p))) 
      p <- length(allele_counts[allele_counts == "A"])/(num_alleles)
      
    
    }
    p_vec[x] <- p
  } 
  return(length(p_vec[p_vec == 0])/length(p_vec))
}

gen_drift_sim2(indiv = 200, p0 = 0.1, n=100, num_sim = 1000)
gen_drift_sim2(indiv = 200, p0 = 0.25, n=100, num_sim = 1000)
gen_drift_sim2(indiv = 200, p0 = 0.5, n=100, num_sim = 1000)
```  
##Question 5  
```{r}
library(reshape)
library(ggplot2)
#Original Genetic drift simulator, only 1 simulation
gen_drift_sim3 <- function(indiv=100, p = 0.5, n=1000)
  {
  num_alleles = indiv * 2
  p_vec <- rep(NA, n)
  p_vec[1] <- p
  
  
  for (i in 2:n){
    allele_counts <- sample(c("A", "a"), size = num_alleles, replace = TRUE, prob = c(p, (1-p))) 
    p <- length(allele_counts[allele_counts == "A"])/(num_alleles)
    p_vec[i] <- p
    
  }
  return(p_vec)
}
gen_drift_sim3()
#Function to assign values of repeated gen_drift simulations into a list. n is #of generations, num_sim is number of simulations
df_drift_sim <- function(indiv=100, p = 0.5, n=100, num_sim = 100) {
  sim_data <- list()
  for (i in 1:num_sim) {
    sims <- unlist(gen_drift_sim3(indiv, p, n))
    sim_data <- rbind(unlist(sim_data), sims)
  } 
 
  sim_data <- as.data.frame(sim_data)
  sim_data <- melt(t(sim_data))
  
} 
sim_data <- df_drift_sim()
sim_data$X1 <- sub("^.", "", sim_data$X1) #Remove V prefix from generation column 
sim_data$X1 <- as.numeric(sim_data$X1) #Convert to numeric
ggplot(sim_data, aes(x = X1, y = value, color = factor(X2))) + 
  geom_line() + 
  labs(x = "Generations", y = "Allele frequency") +
  theme(legend.position="none")
```

##Question 6  

###Function to generate p_value of user-defined linear trend
```{r}
p_val <- function(a, b, num_obs, res_err) {
  x <- seq(from = 1, to = 10, length.out = num_obs)
  y_det <- a + b * x
  y_sim <- rnorm(length(x), mean = y_det, sd = res_err)
  mod_sim <- lm(y_sim ~ x)
  p_val_slope <- summary(mod_sim)$coef[2,4]
  return(p_val_slope)
}
p_val(0.5, 0.1, 20, 2) 
p_val(0.5, 0.1, 20, 2) 
p_val(0.5, 0.1, 20, 2) 
```  
Since this is a stochastic model, rerunning the function multiple times should yield different proportions even with the same inputs which is observed above

###Distribution of p_values over 1000 simulations

```{r}
rep_p_val <- rep(0, 1000)
for (i in 1:length(rep_p_val)) {
  rep_p_val[i] <- p_val(0.5, 0.1, 20, 2)
} #loop 1000 times p-value power analysis
rep_p_val2 <- as.data.frame(unlist(rep_p_val))
colnames(rep_p_val2) <- "p_val"
ggplot(rep_p_val2, aes(x=p_val)) + 
  geom_histogram(bins = 100) +
  labs(x = "p-value") 
```

###Proportion of p-values less than 0.05 

```{r}
length(rep_p_val[rep_p_val<0.05])/length(rep_p_val)
```

###Distribution of p_values if slope is 0

```{r, message = F}
rep_p_val <- rep(0, 1000)
for (i in 1:length(rep_p_val)) {
  rep_p_val[i] <- p_val(0.5, 0, 20, 2)
} #loop 1000 times p-value power analysis
rep_p_val2 <- as.data.frame(unlist(rep_p_val)) #Convert to dataframe
colnames(rep_p_val2) <- "p_val"
ggplot(rep_p_val2, aes(x=p_val)) + 
  geom_histogram(bins = 100) +
  labs(x = "p-value") 
length(rep_p_val[rep_p_val<0.05])/length(rep_p_val)
```  
The counts of p-values are more evenly distributed as opposed to being skewed towards the right (greater proportion of p-values less than 0.5) as seen when the slope = 0.1 compared to when the slope = 0. This is because there is no effect of x on y if the slope = 0 and therefore the p-values will not skew to being less than or greater than 0.5, confirming that the null hypothesis is true and the slope is not equal to zero. 

###Distribution of p_values over 100 simulations throughout different sample sizes

```{r}
sample_sizes <- seq(10, 100, 5) #list of values from 10 to 100 in increments of 5 to act as sample size values
p_freq <- rep(0, length(sample_sizes))
for (i in 1:length(sample_sizes)) {
  p_val_sim_mat <- rep(0, 100) #make empty matrix to hold simulation 
  for (x in 1:100) {
    p_val_sim_mat[x] <- p_val(0.5, 0.1, sample_sizes[i], 1.5)
  } #input p_values into empty matrix 
  p_freq[i] <- length(p_val_sim_mat[p_val_sim_mat < 0.05])/length(p_val_sim_mat)
}
p_freq_sim <- as.data.frame(p_freq)
ggplot(p_freq_sim, aes(x = sample_sizes, y = p_freq)) + geom_bar(stat="identity") +
  labs(x = "Sample size", y = "Frequency of P-values < 0.05")
```

